\textbf{ Accurate spectroscopic estimates of luminosity and distance\\ for luminous red-giant stars }

### 2018-07-24 paper/sp.tex
method section

\item[good features] We assume that the spectrophotometric features are known for
each star with such high fidelity (both precision and accuracy) that we do not
need to account for errors or uncertainties or biases in the features.
That is, we assume that the features are substantially higher in signal-to-noise than the
quantities we are trying to predict.
This is true for most features for most stars, but it does not hole universally.

@@

\item[linearity] Perhaps the most restrictive assumption of this work is that
the stellar logarithmic luminosity (or, equivalently, the absolute magnitude) in
the \zmass\ $K_s$~band can be predicted as a \emph{completely linear function} of
the chosen features. We are only assuming this linearity in a small range of stellar
surface gravity $\logg$, but this assumption is strong, and limits strongly the
flexibility or capacity of the model.

@@

\begin{eqnarray}
Q_n &\equiv& 0.01\,\varpi_n\,10^{0.2\,K_n}
\\
\sigma_n &\equiv& \sigma_{\varpi n}\,10^{0.2\,K_n}
\\
Q_n &=& \theta_Q\cdot x_n + \mbox{noise}
\\
\chi^2 &\equiv& \sum_{n=1}^N \frac{[Q_n - \theta_Q\cdot x_n]^2}{\sigma_n^2}
\\
\hat{\theta}_Q &\leftarrow& \argmin_{\theta_Q}\left[\frac{1}{2}\,\chi^2 + \lambda_Q\,||\theta_Q||_1^1\right]
\quad ,
\end{eqnarray}
where...

\begin{eqnarray}
Q_n &=& \exp(\theta_M\cdot x_n) + \mbox{noise}
\\
M_{Kn} &\approx& \frac{1}{0.2\,\ln 10}\,\theta_M\cdot x_n
\\
\chi^2 &\equiv& \sum_{n=1}^N \frac{[Q_n - \exp(\theta_M\cdot x_n)]^2}{\sigma_n^2}
\\
\hat{\theta}_M &\leftarrow& \argmin_{\theta_M}\left[\frac{1}{2}\,\chi^2 + \lambda_M\,||\theta_M||_1^1\right]
\quad ,
\end{eqnarray}
where...

>>>>

\item[good features] We assume that the spectrophotometric features are known for
each star with such high fidelity (both precision and accuracy) that we do not
need to account for errors or uncertainties or biases in the features.
That is, we assume that the features are substantially higher in signal-to-noise than the
quantities we are trying to predict; in particular we are assuming that the photometry
and the spectroscopy is better measured than the astrometry.
This is true for most features for most stars, but it does not hold universally.

@@

\item[linearity] Perhaps the most restrictive assumption of this work is that
the stellar logarithmic luminosity (or, equivalently, the absolute magnitude) in
the \zmass\ $K_s$~band can be predicted as a \emph{completely linear function} of
the chosen features. We are only assuming this linearity in a small range of stellar
surface gravity $\logg$, but this assumption is strong, and limits strongly the
flexibility or capacity of the model.
We make it to ensure that our method is easy to optimize, and the results are easy
to interpret.

@@

Under the assumptions stated in the previous Section, there are not many choices
for building a model of the stars that is justifiable probabilistically.
Here we lay out the model and methodology, and apply it in the following Section.

Since the photometry is (by the ``good features'' assumption above)
better measured than the parallaxes, we
presume that the parallax uncertainties dominate
the uncertainties in the luminosity measurements.
For each star $n$ in our training set, we define an inverse-square-root
luminosity $Q_n$ estimate that will be subject to Gaussian noise under these assumptions.
\begin{eqnarray}
Q_n &\equiv& \varpi_n\,10^{0.2\,(K_n - 10)}
\\
\sigma_n &\equiv& \sigma_{\varpi n}\,10^{0.2\,(K_n - 10)}
\quad,
\end{eqnarray}
where
$\varpi_n$ is the \gaia\ Catalog entry for the parallax for star $n$,
$K_n$ is the \zmass\ Catalog $K_s$-band magnitude for star $n$,
the $Q_n$ have Gaussian errors of root-variance $\sigma_n$,
and $\sigma_{\varpi n}$ is the \gaia\ Catalog entry for the uncertainty on parallax
for star $n$.
In what follows, the $Q_n$ are the training data.
Naively, the $Q_n$ are related to the absolute magnituges $M_{Kn}$ for the stars by
\begin{eqnarray}
M_{Kn} &\approx& \frac{1}{0.2\,\ln 10}\,\ln Q_n
\quad.
\end{eqnarray}
These would be the absolute magnitudes of the stars under the
``the distance is the inverse of the
measured parallax'' assumption, although we never, in this work, ever make that
assumption, by God.
We construct the $Q_n$ because they are related directly to the absolute magnitudes,
but they (under assumptions) are subject to known Gaussian noise.

The log-likelihood function can be expressed heuristically as
\begin{eqnarray}
Q_n &=& \exp(\theta\cdot x_n) + \mbox{noise}
\\
\chi^2 &\equiv& \sum_{n=1}^N \frac{[Q_n - \exp(\theta\cdot x_n)]^2}{\sigma_n^2}
\quad ,
\end{eqnarray}
where
the model is that the $\ln Q_n$ can be expressed as a linear combination of the components
of some $D$-dimensional feature vector $x_n$,
$\theta$ is a $D$-vector of linear coefficients,
and $\chi^2$ is the negative-log-likelihood
under the assumption of known Gaussian noise and
that there are $N$ independently measured stars $n$.
This feature vector $x_n$ contains photometry in a few bands, and all 8000-ish pixels
in the pseudo-continuum-normalized \apogee\ spectrum, so $D$ is on the order of 8000.

In addition, we assume that many entries in the $D$-vector will be zero
(the ``sparsity'' assumption).
In order to encourage this, we optimize not $\chi^2$ but a regularized objective function
\begin{eqnarray}
\hat{\theta} &\leftarrow& \argmin_{\theta}\left[\frac{1}{2}\,\chi^2 + ||\lambda\cdot\theta||_1^1\right]
\quad ,
\end{eqnarray}
where
$\lambda$ is a $D$-vector of regularization parameters (such that we can regularize
different parts of the $x_n$ vectors differently),
and $||x||_1^1$ is the L1-norm or sum of absolute values of the components of $x$.
This kind of regularization adds a convex term to the optimization and leads to
sparse optima.
We will set the value $\lambda$ of the $D$-vector of regularization strengths by a
very coarse cross-validation.
